Wavelet Multiresolution Spatio-Temporal Modelling Using the Integro-Difference Equation

### 2011-02-23 Multiresolution_framework/ltx/manuscript.tex
deleted extra files

Note that the parameter $\varsigma$ is absorbed into the kernel for simplicity. Oservation equation is given by 

@@

A  two-dimensional MRA can be implemented using 2-D scaling and wavelet functions built up  using the tensor-product approach \cite{Meyer1992,Mallat1998}. We write the 2-D scaling function as
 \phi_{j,\mathbf{l}}\left(r_1,r_2\right)=\phi_{j,l_1}\left(r_1\right)\phi_{j,l_2}\left(r_2\right)
\label{eq:2Dscalingfunction}
and three 2-D wavelet functions $ \psi^{(n)} \left(\mathbf{r}\right)$ where $n=1,2,3$ as
 \psi_{j,\mathbf{l}}^{(1)}\left(r_1,r_2\right)=\phi_{j,l_1}\left(r_1\right)\psi_{j,l_2}\left(r_2\right)
\label{eq:2Dwavelet_1}
 \psi_{j,\mathbf{l}}^{(2)}\left(r_1,r_2\right)=\psi_{j,l_1}\left(r_1\right)\phi_{j,l_2}\left(r_2\right)
\label{eq:2Dwavelet_2}
 \psi_{j,\mathbf{l}}^{(3)}\left(r_1,r_2\right)=\psi_{j,l_1}\left(r_1\right)\psi_{j,l_2}\left(r_2\right)
\label{eq:2Dwavelet_3}
for $j \in \mathbb{Z}$ and, $\mathbf{l} \in \mathbb{Z}^2$, where $\psi_{j,\mathbf{l}}^{(1)},  \psi_{j,\mathbf{l}}^{(2)}$ and $\psi_{j,\mathbf{l}}^{(3)} $ extract fine features of the 2-D field at vertical, horizontal and diagonal orientations respectively.
This can be generalised to $d$-dimensional MRA by introducing the scaling function
\setlength{\arraycolsep}{0.0em}
  \phi_{j,\mathbf{l}}\left(r_1,r_2, \cdots r_d\right)&=&\phi_{j,l_1}\left(r_1\right)\phi_{j,l_2}\left(r_2\right) \cdots \phi_{j,l_d}\left(r_d\right) \label{eq:d-dimensionalscalingfunction}\\
 &&j \in \mathbb{Z}, \quad \mathbf{l} \in \mathbb{Z}^d \nonumber
\setlength{\arraycolsep}{5pt}
and $2^d-1$ wavelet functions $ \psi^{(n)} \left(\mathbf{r}\right), n=1, \cdots,2^d-1$ with a procedure similar to the 2-D case. Using these basis functions, multi-dimensional MRA of the kernel and the field can be obtained by
\setlength{\arraycolsep}{0.0em}
 k\left(\mathbf{r}-\mathbf{r'}\right)&=&\sum_{\mathbf{l}\in \mathbb{Z}^d}\alpha_{j_0,\mathbf{l}}\phi_{j_0,\mathbf{l}}\left(\mathbf{r}-\mathbf{r'}\right) \nonumber \\
&+&\sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l}\in \mathbb{Z}^d}\sum_{i=1}^{2^d-1}\beta_{j,\mathbf{l}}^{(i)} \psi_{j,\mathbf{l}}^{(i)}\left(\mathbf{r}-\mathbf{r'}\right) 
\label{eq:MultidimensionalKernelExpansion}
\setlength{\arraycolsep}{5pt}
z_t\left(\mathbf{r}\right)=\sum_{\mathbf{l}\in \mathbb{Z}^d}x_{t,j_{0},\mathbf{l}}\phi_{j_{0},\mathbf{l}}\left(\mathbf{r}\right)+\sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l}\in \mathbb{Z}^d}\sum_{i=1}^{2^d-1} \check{x}_{t,j,\mathbf{l}}^{(i)}\psi_{j,\mathbf{l}}^{(i)}\left(\mathbf{r}\right)
\label{eq:MultidimensionalFieldExpansion}
Equations (\ref{eq:MultidimensionalKernelExpansion}) and (\ref{eq:MultidimensionalFieldExpansion}) are infinite series expansions. For practical implementation, (\ref{eq:MultidimensionalKernelExpansion}) and (\ref{eq:MultidimensionalFieldExpansion}) must be truncated at some level $j$, therefore we have

@@

Each vector of basis functions on the right hand-side of (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in a way similar to their corresponding weight vectors in (\ref{KernelWeights}) and (\ref{FieldWeights}). Note that the identified IDE model will approximate the underlying spatio-temporal dynamics at different spatial scales. In this way both macroscopic and microscopic behaviour of the system can be represented simultaneously.

To derive the state-space representation of the model we substitute decompositions \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} into equation \eqref{DiscreteTimeModel} and then cross-multiply the decomposed model by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the spatial domain, $\Omega$, giving
\begin{align}\label{eq:DecomposedModel2} 
	&\int_{\Omega} \boldsymbol\mu  \left(\mathbf{r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^\top  d\mathbf{r}\mathbf{x}_{t+1}= 
	\xi\int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^\top d\mathbf{r}\mathbf{x}_t \nonumber \\
	&+T_s \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r'}\right)
	    \boldsymbol\mu^\top\left(\mathbf{r'}\right)\mathbf{x}_t 
	\, d\mathbf{r}'d\mathbf{r}}
	+ \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}, 
\end{align}
Now defning 
\begin{equation}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} \triangleq \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\boldsymbol\mu^\top\left(\mathbf{r}\right) d\mathbf{r}
\end{equation}
\begin{equation}\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta}\triangleq\int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r'}\right)
	    \boldsymbol\mu^\top\left(\mathbf{r'}\right)\mathbf{x}_t 
	\, d\mathbf{r}'d\mathbf{r}}
\end{equation}
and substituting these into \eqref{eq:DecomposedModel2} and cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ yields
\begin{equation}\label{eq:StateEquation}
 \mathbf x_{t+1}=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t
\end{equation}
where
\begin{equation}\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta)=(T_s\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I)
\end{equation}
and $\mathbf I$ is the identity matrix, and 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance \cite{Scerri2009}
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\eta\left(\mathbf{r}-\mathbf{r'}\right)\boldsymbol\mu^{\top}\left(\mathbf{r'}\right)d\mathbf{r'}d\mathbf{r}\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
Similarly, the observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is 
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\mu}_j(\mathbf{r}') \, d\mathbf{r}'.
\end{equation}
This completes the state-space representation of the multiresolution approximation of the neural field model. 
\parham{Mike you can add state-space derivation using Galerkin projection here.}
\section{B-spline wavelet IDE model}
Multiresolution approximation can be implemented using different classes of scaling and wavelet functions. In order to determine the matrices $ \boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$ convolution and inner product expressions of the basis functions need to be computable. B-spline wavelet and scaling functions are particularly convenient in the above decomposition as their convolution and inner product expressions can be found analytically.

The  $m$th order cardinal B-spline scaling function is defined by the recurrence relation \cite{Chui1992} 
\begin{eqnarray}
N_{m}\left(s\right)&=&\left(N_{m-1}\ast N_{1}\right)\left(s\right)\nonumber \\
&=&\int_0^{1} N_{m-1}\left( s-r\right)dr \quad m>1 
\label{SplineConvolutionIntegral}
\end{eqnarray}
Here $\ast$ denotes the convolution operator and $N_1\left(s\right)$ is the characteristic function of the unit interval  $\left[ 0,1\right) $
\begin{equation}
N_{1}\left(s\right)=
\begin{cases}
1 & \text{if $ 0\le s<1$}, \\
0 & \mathrm{elsewhere}
\end{cases}
\end{equation}
In order to compute the  integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2}) we need to be able to calculate the convolution and inner product of B-spline scaling functions. Equation (\ref{SplineConvolutionIntegral}) can be rewritten as $(m-1)$ times convolution of indicator function with itself 
\begin{equation}
 N_{m}\left(s\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(s\right)
\end{equation}
then by using the associativity property of the convolution operator, it is clear that
\setlength{\arraycolsep}{0.0em}
\begin{eqnarray}
N_{m}\left(s\right) \ast N_{m'}\left(s\right)&=&\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(s\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(s\right)\nonumber\\
&=&N_{m+m'}\left(s\right)
\label{eq:BsplineConvolution}
\end{eqnarray}
\setlength{\arraycolsep}{5pt}
This result can be used to show that (Lemma \ref{lemma:InProduct}, Appendix)
\setlength{\arraycolsep}{0.0em}
\begin{eqnarray}
 \left\langle N_{m}\left(s-l_{1}\right), N_{m'}\left(s-l_{2}\right)\right\rangle&=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
&=&N_{m+m'}\left(m'+l_{2}-l_{1}\right)
\label{eq:BsplineInnerProduct}
\end{eqnarray}
\setlength{\arraycolsep}{5pt}
where $\left\langle  .,.\right\rangle $ denotes the inner product.

@@

Where the subscript $m$ is introduced to $\phi_{j,l}(s)$ and $\psi_{j,l}(s) $ to indicate the $m$th order B-spline scaling and wavelet functions. In the two dimensional case, the convolution integral can be factorised as follows, e.g. 
\setlength{\arraycolsep}{0.14em}
\begin{IEEEeqnarray}{!l l }
 \phi_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right)\ast \psi^{1}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)&=\left[\phi_{m;j,l_1}\left(s_{1}\right)\ast\phi_{m;j,l'_1}\left(s_{1}\right)\right] \nonumber \\
&\times \left[ \phi_{m;j,l_2}\left(s_{2}\right) \ast \psi_{m;j,l'_2}\left(s_{2}\right)\right] \nonumber \\
&
\end{IEEEeqnarray}
\setlength{\arraycolsep}{5pt}
The preceding factorisation can be also applied to the 2D-inner product integral. Non-zero quantities can  be then expressed by
\setlength{\arraycolsep}{0.14em}
\begin{IEEEeqnarray}{!l l }
\left\langle \phi_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\phi_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle &=\left\langle \phi_{m;j,l_1}\left(s_{1}\right),\phi_{m;j,l'_1}\left(s_{1}\right)\right\rangle \nonumber \\
& \times\left\langle \phi_{m;j,l_2}\left(s_{2}\right) , \phi_{m;j,l'_2}\left(s_{2}\right)\right\rangle \nonumber \\
&
\end{IEEEeqnarray}
\setlength{\arraycolsep}{5pt}
likewise we can compute 
$\left\langle  \psi^{1}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\psi^{1}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle$
, \\
$\left\langle  \psi^{2}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\psi^{2}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle$ and \\
$
\left\langle \psi^{3}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right), \psi^{3}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle $. \\ \\
Using these expansions it becomes straightforward to calculate the integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2}). In this work, 4th order cardinal B-spline scaling and wavelet functions are used therefore, to compute the integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2})  the $8$th and $12$th order B-spline values at integer points are  required and can be found in \cite{Goswami1999}.

@@

In the E-step, given an estimate of $\boldsymbol\theta$, the Rauch Tung Striebel (RTS) smoother can be employed to determine the posterior state sequence. It can be shown that \cite{Dempster1977,Gibsona2005} every EM iteration increases the lower bound on the likelihood function and hence improving the state and parameter estimates. The stopping criterion is usually associated  with  either the change in the parameter estimates or the log-likelihood variation \cite{McLachlan1997}. The stopping rule adopted herein is
\left(\parallel \mathbf{A} \parallel_{F}^{(i)}-\parallel \mathbf{A} \parallel_{F}^{(i-1)}\right)<\varepsilon 
where $\varepsilon$ is a threshold value and $\parallel \mathbf{A} \parallel_{F}^{(i)}$ and $ \parallel \mathbf{A} \parallel_{F}^{(i-1)}$ are the Frobenius norm  of the successive estimates of $\mathbf{A} $ matrices. It has been shown \cite{Dewar2009} that the state dimension is a critical factor in the complexity of the estimation procedure, however using the above decomposition, this is decoupled from the number of observation locations.

@@

\setlength{\arraycolsep}{0.0em}
k\left(s-r\right)&=& -\phi_{4;0,-3}\left(s-r\right)+\phi_{4;0,-2}\left(s-r\right)\nonumber\\
&&{-}\:0.5\psi_{4;1,-6}\left(s-r\right)+0.5\psi_{4;1,-3}\left(s-r\right)
\label{eq:1DSimulationKernel}
\setlength{\arraycolsep}{5pt}

@@

 \caption{The average (over time) power in dB of the spatial frequency of the observations. The dashed line shows the cutoff frequency (-3 dB point).}

@@

 \caption{Spatial field at one time instant for different spatial resolutions; estimated and true fields are
shown by dashed and solid lines respectively.}

@@

\begin{table}[!h]
\renewcommand{\arraystretch}{1.3}
\caption {MISE  at particular scale between measured and estimated 2D-field  at a particular time}
\label{table:ValidationResultField}
\begin{tabular}{cccc}
\hline \hline
\multicolumn{2}{r}{Time}&\multicolumn{2}{c}{MISE} \\
\cline{3-4}
&&$j=0$&$j=1$ \\ 
\hline 
&$5$ &$5.424$&$0.609$ \\
&$20$ &$5.422$&$0.596$ \\
&$50$ &$5.519$&$0.626$ \\
&$100$ &$5.488$&$0.611$ \\
&$150$ &$5.494$&$0.587$ \\
\hline \hline
\end{tabular}
\end {table}
\section{Conclusion}
A multi-resolution approach to modelling spatio-temporal systems has been presented. This model is able to represent continuous-space, discrete-time dynamics at a number of spatial scales simultaneously. This ability greatly extends the class of system which the Integro-Difference Equation can represent.

By decomposing both the spatial field and the spatial mixing kernel using a wavelet decomposition, it becomes possible to represent small scale details in the field at the same time as large scale details. This is an essential component of any practical spatio-temporal model, without which the model must be artificially separated into global and local modes a-priori.

>>>>

Note that the parameter $\varsigma$ is absorbed into the kernel for simplicity. Observation equation is given by 

@@

% A  two-dimensional MRA can be implemented using 2-D scaling and wavelet functions built up  using the tensor-product approach \cite{Meyer1992,Mallat1998}. We write the 2-D scaling function as
% \begin{equation}
%  \phi_{j,\mathbf{l}}\left(r_1,r_2\right)=\phi_{j,l_1}\left(r_1\right)\phi_{j,l_2}\left(r_2\right)
% \label{eq:2Dscalingfunction}
% \end{equation}
% and three 2-D wavelet functions $ \psi^{(n)} \left(\mathbf{r}\right)$ where $n=1,2,3$ as
% \begin{equation}
%  \psi_{j,\mathbf{l}}^{(1)}\left(r_1,r_2\right)=\phi_{j,l_1}\left(r_1\right)\psi_{j,l_2}\left(r_2\right)
% \label{eq:2Dwavelet_1}
% \end{equation}
% \begin{equation}
%  \psi_{j,\mathbf{l}}^{(2)}\left(r_1,r_2\right)=\psi_{j,l_1}\left(r_1\right)\phi_{j,l_2}\left(r_2\right)
% \label{eq:2Dwavelet_2}
% \end{equation}
% \begin{equation}
%  \psi_{j,\mathbf{l}}^{(3)}\left(r_1,r_2\right)=\psi_{j,l_1}\left(r_1\right)\psi_{j,l_2}\left(r_2\right)
% \label{eq:2Dwavelet_3}
% \end{equation}
% for $j \in \mathbb{Z}$ and, $\mathbf{l} \in \mathbb{Z}^2$, where $\psi_{j,\mathbf{l}}^{(1)},  \psi_{j,\mathbf{l}}^{(2)}$ and $\psi_{j,\mathbf{l}}^{(3)} $ extract fine features of the 2-D field at vertical, horizontal and diagonal orientations respectively.
% This can be generalised to $d$-dimensional MRA by introducing the scaling function
% \setlength{\arraycolsep}{0.0em}
% \begin{eqnarray}
%   \phi_{j,\mathbf{l}}\left(r_1,r_2, \cdots r_d\right)&=&\phi_{j,l_1}\left(r_1\right)\phi_{j,l_2}\left(r_2\right) \cdots \phi_{j,l_d}\left(r_d\right) \label{eq:d-dimensionalscalingfunction}\\
%  &&j \in \mathbb{Z}, \quad \mathbf{l} \in \mathbb{Z}^d \nonumber
% \end{eqnarray}
% \setlength{\arraycolsep}{5pt}
% and $2^d-1$ wavelet functions $ \psi^{(n)} \left(\mathbf{r}\right), n=1, \cdots,2^d-1$ with a procedure similar to the 2-D case. Using these basis functions, multi-dimensional MRA of the kernel and the field can be obtained by
% \setlength{\arraycolsep}{0.0em}
% \begin{eqnarray}
%  k\left(\mathbf{r}-\mathbf{r'}\right)&=&\sum_{\mathbf{l}\in \mathbb{Z}^d}\alpha_{j_0,\mathbf{l}}\phi_{j_0,\mathbf{l}}\left(\mathbf{r}-\mathbf{r'}\right) \nonumber \\
% &+&\sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l}\in \mathbb{Z}^d}\sum_{i=1}^{2^d-1}\beta_{j,\mathbf{l}}^{(i)} \psi_{j,\mathbf{l}}^{(i)}\left(\mathbf{r}-\mathbf{r'}\right) 
% \label{eq:MultidimensionalKernelExpansion}
% \end{eqnarray}
% \setlength{\arraycolsep}{5pt}
% \begin{equation}
% z_t\left(\mathbf{r}\right)=\sum_{\mathbf{l}\in \mathbb{Z}^d}x_{t,j_{0},\mathbf{l}}\phi_{j_{0},\mathbf{l}}\left(\mathbf{r}\right)+\sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l}\in \mathbb{Z}^d}\sum_{i=1}^{2^d-1} \check{x}_{t,j,\mathbf{l}}^{(i)}\psi_{j,\mathbf{l}}^{(i)}\left(\mathbf{r}\right)
% \label{eq:MultidimensionalFieldExpansion}
% \end{equation}
Equations \eqref{eq:KernelExpansion} and \eqref{eq:FieldExpansion} are infinite series expansions. For practical implementation,  \eqref{eq:KernelExpansion} and \eqref{eq:FieldExpansion} must be truncated at some level $j$, therefore we have

@@

Each vector of basis functions on the right hand-side of (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in a way similar to their corresponding weight vectors in (\ref{KernelWeights}) and (\ref{FieldWeights}). 
% Note that the identified IDE model will approximate the underlying spatio-temporal dynamics at different spatial scales. In this way both macroscopic and microscopic behaviour of the system can be represented simultaneously.

To derive the state-space representation of the model we substitute decompositions \eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} into equation \eqref{DiscreteTimeModel} and then cross-multiply the decomposed model by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the spatial domain, $\Omega$, giving
\begin{align}\label{eq:DecomposedModel2} 
	&\int_{\Omega} \boldsymbol\mu  \left(\mathbf{r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^\top  d\mathbf{r}\mathbf{x}_{t+1}= 
	\xi\int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^\top d\mathbf{r}\mathbf{x}_t \nonumber \\
	&+T_s \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r'}\right)
	    \boldsymbol\mu^\top\left(\mathbf{r'}\right)\mathbf{x}_t 
	\, d\mathbf{r}'d\mathbf{r}}
	+ \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}, 
\end{align}
Now defning 
\begin{equation}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} \triangleq \int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\boldsymbol\mu^\top\left(\mathbf{r}\right) d\mathbf{r}
\end{equation}
\begin{equation}\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta}\triangleq\int_{\Omega}\boldsymbol\mu\left(\mathbf{r}\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(\mathbf{r}-\mathbf{r'}\right)
	    \boldsymbol\mu^\top\left(\mathbf{r'}\right)\mathbf{x}_t 
	\, d\mathbf{r}'d\mathbf{r}}
\end{equation}
and substituting these into \eqref{eq:DecomposedModel2} and cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ yields
\begin{equation}\label{eq:StateEquation}
 \mathbf x_{t+1}=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t
\end{equation}
where
\begin{equation}\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta)=(T_s\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I)
\end{equation}
and $\mathbf I$ is the identity matrix, and 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance \cite{Scerri2009}
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(\mathbf{r}\right)\eta\left(\mathbf{r}-\mathbf{r'}\right)\boldsymbol\mu^{\top}\left(\mathbf{r'}\right)d\mathbf{r'}d\mathbf{r}\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
Similarly, the observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is 
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\mu}_j(\mathbf{r}') \, d\mathbf{r}'.
\end{equation}
This completes the state-space representation of the multiresolution approximation of the neural field model. 
% \parham{Mike you can add state-space derivation using Galerkin projection here.}
\section{B-spline wavelet IDE model}
Multiresolution approximation can be implemented using different classes of scaling and wavelet functions. In order to determine the matrices $ \boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$ convolution and inner product expressions of the basis functions need to be computable. B-spline wavelet and scaling functions are particularly convenient in the above decomposition as their convolution and inner product expressions can be found analytically.

The  $m$th order cardinal B-spline scaling function is defined by the recurrence relation \cite{Chui1992} 
\begin{eqnarray}
N_{m}\left(s\right)&=&\left(N_{m-1}\ast N_{1}\right)\left(s\right)\nonumber \\
&=&\int_0^{1} N_{m-1}\left( s-r\right)dr \quad m>1 
\label{SplineConvolutionIntegral}
\end{eqnarray}
Here $\ast$ denotes the convolution operator and $N_1\left(s\right)$ is the characteristic function of the unit interval  $\left[ 0,1\right) $
\begin{equation}
N_{1}\left(s\right)=
\begin{cases}
1 & \text{if $ 0\le s<1$}, \\
0 & \mathrm{elsewhere}
\end{cases}
\end{equation}
In order to compute the  integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2}) we need to be able to calculate the convolution and inner product of B-spline scaling functions. Equation (\ref{SplineConvolutionIntegral}) can be rewritten as $(m-1)$ times convolution of indicator function with itself 
\begin{equation}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right)
\end{equation}
then by using the associativity property of the convolution operator, it is clear that
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right)
\end{align}
A direct consequence of \eqref{eq:BsplineConvolution} is 
\begin{eqnarray}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle&=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
&=&N_{m+m'}\left(m'+l_{2}-l_{1}\right)
\label{eq:BsplineInnerProduct}
\end{eqnarray}
where $\left\langle  .,.\right\rangle $ denotes the inner product.

@@

% Where the subscript $m$ is introduced to $\phi_{j,l}(s)$ and $\psi_{j,l}(s) $ to indicate the $m$th order B-spline scaling and wavelet functions. In the two dimensional case, the convolution integral can be factorised as follows, e.g. 
% \setlength{\arraycolsep}{0.14em}
% \begin{IEEEeqnarray}{!l l }
%  \phi_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right)\ast \psi^{1}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)&=\left[\phi_{m;j,l_1}\left(s_{1}\right)\ast\phi_{m;j,l'_1}\left(s_{1}\right)\right] \nonumber \\
% &\times \left[ \phi_{m;j,l_2}\left(s_{2}\right) \ast \psi_{m;j,l'_2}\left(s_{2}\right)\right] \nonumber \\
% &
% \end{IEEEeqnarray}
% \setlength{\arraycolsep}{5pt}
% The preceding factorisation can be also applied to the 2D-inner product integral. Non-zero quantities can  be then expressed by
% \setlength{\arraycolsep}{0.14em}
% \begin{IEEEeqnarray}{!l l }
% \left\langle \phi_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\phi_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle &=\left\langle \phi_{m;j,l_1}\left(s_{1}\right),\phi_{m;j,l'_1}\left(s_{1}\right)\right\rangle \nonumber \\
% & \times\left\langle \phi_{m;j,l_2}\left(s_{2}\right) , \phi_{m;j,l'_2}\left(s_{2}\right)\right\rangle \nonumber \\
% &
% \end{IEEEeqnarray}
% \setlength{\arraycolsep}{5pt}
% likewise we can compute 
% $\left\langle  \psi^{1}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\psi^{1}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle$
% , \\
% $\left\langle  \psi^{2}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\psi^{2}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle$ and \\
% $
% \left\langle \psi^{3}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right), \psi^{3}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle $. \\ \\
In this work, 4th order cardinal B-spline scaling and wavelet functions are used therefore, to compute the integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2})  the $8$th and $12$th order B-spline values at integer points are  required and can be found in \cite{Goswami1999}.

@@

In the E-step, given an estimate of $\boldsymbol\theta$, the Rauch Tung Striebel (RTS) smoother can be employed to determine the posterior state sequence. It can be shown that \cite{Dempster1977,Gibsona2005} every EM iteration increases the lower bound on the likelihood function and hence improving the state and parameter estimates.

@@

% \setlength{\arraycolsep}{0.0em}
% \begin{eqnarray}
% k\left(s-r\right)&=& -\phi_{4;0,-3}\left(s-r\right)+\phi_{4;0,-2}\left(s-r\right)\nonumber\\
% &&{-}\:0.5\psi_{4;1,-6}\left(s-r\right)+0.5\psi_{4;1,-3}\left(s-r\right)
% \label{eq:1DSimulationKernel}
% \end{eqnarray}
% \setlength{\arraycolsep}{5pt}

@@

 \caption{(a) Mexican-hat connectivity kernel. The kernel is a composite of two B-spline
functions (dashed lines), representing local excitation and lateral inhibition. (b) The average (over time) power in dB of the spatial frequency of the observations generated by the kernel in (a). The dashed line shows the cutoff frequency (-3 dB point).}

@@

 \caption{Spatial field at one time instant for different spatial resolutions, estimated and true fields are
shown by dashed and solid lines respectively; (a) $j=0$, $n_x=17$. (b) $j=1$, $n_x=33$. (c) $j=2$, $n_x=65$. (d) $j=3$, $n_x=131$. (e) $j=4$, $n_x=263$.}

@@

% \begin{table}[!h]
% \renewcommand{\arraystretch}{1.3}
% \caption {MISE  at particular scale between measured and estimated 2D-field  at a particular time}
% \label{table:ValidationResultField}
% \begin{tabular}{cccc}
% \hline \hline
% \multicolumn{2}{r}{Time}&\multicolumn{2}{c}{MISE} \\
% \cline{3-4}
% &&$j=0$&$j=1$ \\ 
% \hline 
% &$5$ &$5.424$&$0.609$ \\
% &$20$ &$5.422$&$0.596$ \\
% &$50$ &$5.519$&$0.626$ \\
% &$100$ &$5.488$&$0.611$ \\
% &$150$ &$5.494$&$0.587$ \\
% \hline \hline
% \end{tabular}
% \end {table}
% \section{Conclusion}
% A multi-resolution approach to modelling spatio-temporal systems has been presented. This model is able to represent continuous-space, discrete-time dynamics at a number of spatial scales simultaneously. This ability greatly extends the class of system which the Integro-Difference Equation can represent.
% 
% 
% By decomposing both the spatial field and the spatial mixing kernel using a wavelet decomposition, it becomes possible to represent small scale details in the field at the same time as large scale details. This is an essential component of any practical spatio-temporal model, without which the model must be artificially separated into global and local modes a-priori.
% 
% 
% The complexity of the model is not affected by the spatial resolution of the observation process, rather it reflects the underlying complexity of the system under study. Therefore, increasing the resolution at which the system is observed does not necessarily increase the complexity of the system identification problem. However the developed algorithm is sensitive to the state dimension, equivalent to the detail represented in the field, and hence intelligent approaches to reduce the number of basis functions is noted as future work. 
% 
